Switch equivalence and weak prime domains for fusions

### 2024-04-22 fusione.tex
typos
minori

@@

\begin{corollary}\label{cor:slice}
	Let  $X$ be an object of an $\mathcal{M}$-adhesive category $\X$. Then  $\X/X$ and $X/\X$ are, respectively, is $\mathcal{M}/X$- and $X/\mathcal{M}$-adhesive, where
	\[\mathcal{M}/X:=\{m\in  \mathcal{A}(\X/X) \mid m\in \mathcal{M} \} \qquad X/\mathcal{M}:=\{m\in  \mathcal{A}(X/\X) \mid m\in \mathcal{M} \}\]
\end{corollary}

@@

We end this section proving two properties of $\mathcal{M}$-adhesive categories:  $\mathcal{M}$-pushout-pullback decomposition and uniqueness of pushouts complements.

@@

Let us turn our attention to pushout complements.

\begin{definition}[Pushout complement]
Let $f\colon X\to Y$ and $g\colon Y\to Z$ ba two composable arrows in a category $\X$. A \emph{pushout complement} for the pair $(f,g)$ is a pair $(h,k)$ with $h\colon X\to W$ and $k\colon W\to Z$ such that the square below commutes and it is a pushout.
\[\xymatrix{X \ar[r]^{f} \ar[d]_{h}& Y \ar[d]^{g} \\ W \ar[r]_{k}& Z}\]
\end{definition}

@@

Working in an $\mathcal{M}$-adhesive category we can amend the second defect.

\begin{lemma}\label{lem:radj}
Let $f\colon X\to Y$ be an arrow in an $\mathcal{M}$-adhesive category $\X$ and suppose that the left square below is a pushout while the  right one is a pullback, with $ m\colon M\to X$ and $n\colon N\to Y$ in $\mathcal{M}$.
\[\xymatrix{M \ar@{>->}[d]_{m}\ar[r]^{h}& Q \ar@{>->}[d]^{k} & P \ar[r]^{p_1} \ar@{>->}[d]_{p_2}& N \ar@{>->}[d]^{n}\\X\ar[r]_{f}&Y &X \ar[r]_f&Y\\}\]
Then $n\leq k$ if and only if $p_2\leq m$.
\end{lemma}
\begin{proof}  
	We start pulling back $n$ along $k$ and then pulling back the resulting arrow along $h$, getting the following two pullbacks square:
\[\xymatrix{B \ar@{>->}[d]_{b}\ar@{>->}[r]^{s}& N \ar@{>->}[d]^{n} & A \ar[r]^{r} \ar@{>->}[d]_{a}& B \ar@{>->}[d]^{b}\\Q\ar@{>->}[r]_{k}&Y &M \ar[r]_h&Q\\}\]
Notice that 
\begin{align*}
	f\circ m\circ a&=k\circ h\circ a\\&=k\circ b\circ r\\&=n\circ s\circ r
\end{align*}
	Thus there exists a unique $t\colon A\to P$ as in the diagram
\[\xymatrix{A \ar@{>->}[d]_{a} \ar@{.>}[dr]^{t}\ar[r]^{r} &B \ar@{>->}@/^.2cm/[dr]^{s} \\ M \ar@{>->}@/_.2cm/[dr]_{m}&P\ar[r]^{p_1}  \ar@{>->}[d]_{p_2}& N \ar@{>->}[d]^{n}\\ &X \ar[r]_{f} & Y}\]
	We can then consider the cube below, in which the front, right and back faces are pullbacks.
		\[\xymatrix@C=15pt@R=15pt{&A\ar@{>->}[dd]|\hole_(.65){a}\ar[rr]^{r} \ar[dl]_{t} && B \ar@{>->}[dd]^{b} \ar[dl]_{s} \\ P  \ar@{>->}[dd]_{p_2}\ar[rr]^(.7){p_1} & & N \ar@{>->}[dd]_(.3){n}\\&M\ar[rr]|\hole^(.65){h} \ar@{>->}[dl]^{m} && Q \ar@{>->}[dl]^{k} \\X \ar[rr]_{f} & & Y}\]
	We can notice that the left square is a pullback too. To see this it is enough to apply \Cref{lem:pb1} to the diagram
	\[\xymatrix{A  \ar@/^.4cm/[rr]^{s\circ r}\ar@{>->}[d]_{a} \ar[r]_{t}& \ar[r]_{p_1} P \ar@{>->}[d]^{p_2}& N \ar[d]^{c}\\ M \ar@/_.4cm/[rr]_{k\circ h}\ar@{>->}[r]^{m}& X \ar[r]^{f}& Y}\]

@@

\begin{corollary}[Uniqueness of pushouts complements]\label{lem:pocomp} 
Let $\X$ be a $\mathcal{M}$-adhesive category. Given $m\colon X\mto Y$ in $\mathcal{M}$ and $f\colon Y\to Z$, let $(h_1, k_1)$ and $(h_2, k_2)$ be pushout complements of $m$  and $f$ and $W_1=\cod{h_1}$, $W_2=\cod{h_2}$. Then there exists a unique isomorphism $\phi\colon W_1\to W_2$ making the following diagram commutative.
\[\xymatrix{&X \ar@{>->}[r]^{m} \ar[d]_{h_1} \ar@/_.3cm/[ddl]_{h_2}& Y \ar[d]^{f} \\ &W_1 \ar@{.>}[dl]^{\phi} \ar@{>->}[r]_{k_1} & Z \\ W_2 \ar@{>->}@/_.3cm/[urr]_{k_2}}\]

@@

\subsubsection{Final pullback complements}

@@

\begin{lemma}\label{lem:pushandpullpb}
	Let $\X$ be a category and suppose that the diagram below is given.
	\[\xymatrix{ P \ar[r]^{p_0} \ar[d]_{p_1}& A  \ar@/^.2cm/[ddr]^{k_0}\ar[d]^{f}\\  B  \ar@/_.2cm/[drr]_{k_1} \ar[r]_g & C \ar[dr]^{h}\\&& D }\]
	Then the following hold true:
	\begin{enumerate}
		\item if the outer boundary is a pullback, then the inner square is a pullback too;
		\item if the inner square is a pullback and $h$ is mono then also the outer square is a pullback.
	\end{enumerate}
\end{lemma}
\begin{proof}\begin{enumerate}
		\item Let $e_0\colon E\to A$ and $e_1\colon E\to B$ be a pair of arrows such that 
		\[f\circ e_0=g\circ e_1\]
		Then we also have 
		\begin{align*}
		k_0\circ e_0 &= h\circ f\circ e_0\\&=h\circ g \circ e_1\\&=k_1\circ e_1
		\end{align*}
		Thus there exists a unique arrow $e\colon E\to P$ such that
		\[p_0\circ e= e_0 \qquad p_1\circ e=e_1\]
		which entails our claim.
		\item We proceed as before considering two arrows $e_0\colon E\to A$ and $e_1\colon E\to B$, but in this case we suppose that 
		\[k_0\circ e_0=k_1\circ e_1\]
		In particular, we have
		\begin{align*}
			h\circ f \circ e_0 &= k_0\circ e_0\\&=k_1\circ e_1\\&=h\circ g\circ e_1
		\end{align*}
		Thus, since $h$ is mono, we can deduce that $f\circ e_0$ and $g\circ e_1$ are equal and the thesis now follows immediately.
		\qedhere 
	\end{enumerate}
	\end{proof} 
	In particular, we will use the first point of the previous lemma to deduce the following result.

	\begin{corollary}\label{cor:cond} Let $f\colon A\to C$, $g\colon B\to C$ and $h\colon C\to D$ be arrows in a category $\X$, let also $l_0\colon A'\to A$, $l_1\colon B'\to B$ be two monomorphisms. Suppose moreover, that the following squares are pullbacks:
		\[\xymatrix@C=35pt{Q \ar[r]^{q_0} \ar[d]_{q_1} & A' \ar[d]^{h\circ f\circ l_0} & P \ar[r]^{p_0} \ar[d]_{p_1} & A \ar[d]^{f} & Q  \ar[r]^{l_0\circ q_0} \ar[d]_{l_1\circ q_1}& A \ar[d]^{h\circ f}\\ B \ar[r]_{h\circ g\circ l_1} &D& B \ar[r]_{g}& C &  B \ar[r]_{h\circ g} & D}\]
		Then there exists an arrow $\phi\colon P\to Q$ such that $(P, \{q_i\circ \phi\}_{i=0}^1)$ is a pullback for $f\circ l_0$ along $g\circ l_1$.
	\end{corollary} 
	\begin{proof}
To prove the existence of $\phi\colon P\to Q$ it's enough to consider the solid part of the diagram below.
\[\xymatrix@C=30pt{P \ar@{.>}[dr]^{\phi} \ar@/^.2cm/[drr]^{p_0} \ar@/_.2cm/[ddr]_{p_1}\\ & Q  \ar[r]^(.4){l_0\circ q_0} \ar[d]^{l_1\circ q_1}& A \ar[d]^{h\circ f} \\ &B \ar[r]_{h\circ g} & D}\]
Notice that $\phi$ is a monomorphism. If $\phi\circ a=\phi \circ b$ for some $a,b\colon T\rightrightarrows P$ then
\[\begin{split}
	p_0\circ a&=l_0\circ q_0\circ \phi \circ a\\&=l_0\circ q_0\circ \phi \circ b\\&=p_0\circ b
\end{split} \qquad \begin{split}
p_1\circ a&=l_1\circ q_1\circ \phi \circ a\\&=l_1\circ q_1\circ \phi \circ b\\&=p_1\circ b
\end{split} \]

>>>>

\begin{remark}\label{rem:pb}
	Let $X$ be an object of $\X$, with $\X$ $\mathcal{M}$-adhesive. We can consider  the full subcategory $\mathcal{M}/X$ of $\X/X$ made by arrows in $\mathcal{M}$. Since $\X$ has all $\mathcal{M}$-pullbacks, for every arrow $f\colon X\to Y$ we have a functor $f^*\colon \mathcal{M}/Y\to \mathcal{M}/X $ sending $m\colon \mto X$ to a chosen pullback of it along $f$.
	\end{remark}

@@

\begin{corollary}\label{cor:slice}
	Let  $X$ be an object of an $\mathcal{M}$-adhesive category $\X$. Then  $\X/X$ and $X/\X$ are, respectively, $\mathcal{M}_X$- and $\mathcal{M}^X$-adhesive, where
	\[\mathcal{M}_X:=\{m\in  \mathcal{A}(\X/X) \mid m\in \mathcal{M} \} \qquad \mathcal{M}^X:=\{m\in  \mathcal{A}(X/\X) \mid m\in \mathcal{M} \}\]
\end{corollary}

@@

We end this section proving a useful property of $\mathcal{M}$-adhesive categories:  $\mathcal{M}$-pushout-pullback decomposition.

@@

\subsubsection{Pushout complements}

We turn now to the exam of the notion of \emph{pushout complement}. We will prove that, in an $\mathcal{M}$-adhesive category $\X$ certain pushout complements are unique and are \emph{final pullbacks complements} in the sense of \cite{dyckhoff1987exponentiable,corradini2006sesqui}.

\begin{definition}[Pushout complement]
Let $f\colon X\to Y$ and $g\colon Y\to Z$ be two composable arrows in a category $\X$. A \emph{pushout complement} for the composable pair $(f,g)$ is a pair $(h,k)$ with $h\colon X\to W$ and $k\colon W\to Z$ such that the square below commutes and it is a pushout.
\[\xymatrix{X \ar[r]^{f} \ar[d]_{h}& Y \ar[d]^{g} \\ W \ar[r]_{k}& Z}\]
\end{definition}

@@

In an $\mathcal{M}$-adhesive category we can approach pushout complements in a more abstract way, using the pullback functor of \Cref{rem:pb}.

\begin{definition}
Let $F\colon \X \to \Y$ be a functor and $Y$ an object of $\Y$. We say that $F$ has a \emph{partial right adjoint at $Y$} if there exists an object $G(Y)$ in $\X$ and an arrow $\epsilon_Y\colon F(G(Y))\to Y$ such that, for every  $f\colon F(X)\to Y$ there exists a unique $g\colon X\to G(Y)$ making the diagram below commutative.
\[\xymatrix@C=15pt{F(X) \ar[rr]^{F(g)} \ar[dr]_{f}&&F(G(Y)) \ar[dl]^{\epsilon_Y}\\ & Y}\]
\end{definition}

\begin{remark}
	It is immediate to see that $F$ is a left adjoint if and only if it has a partial right adjoint at every $Y\in \Y$.
\end{remark}

We can prove two  useful properties of partial right adjoints.

\begin{proposition}\label{prop:uni}
	Let $F\colon \X\to \Y$ be a functor and suppose that it has a partial right adjoint $(G(Y), \epsilon_Y)$ at $Y$. Then the following hold true:
	\begin{enumerate}
\item there is a natural isomorphism $\psi\colon \X(-, G(Y))\to \Y(F(-), Y)$;
\item if $(G'(Y), \epsilon'_Y)$ is another partial right adjoint to $F$ at $Y$, then there exists a unique isomorphism $\phi_Y\colon G(Y)\to G'(Y)$ such that the diagram below commutes.
\[\xymatrix@C=15pt{F(G(Y)) \ar[rr]^{F(\phi_Y)} \ar[dr]_{\epsilon_Y} & &F(G'(Y)) \ar[dl]^{\epsilon'_Y}\\ & Y}\]
	\end{enumerate}
\end{proposition}
\begin{proof}
	\begin{enumerate}
		\item Given an object $X$ of $\X$ we can define $\phi_X$ putting
		\begin{align*}
			\phi_X \colon \X(X, G(Y))&\to \Y(F(X), Y)\\
			g&\mapsto \epsilon_Y\circ F(g)
		\end{align*}
		By definition of partial right adjoint $\phi_X$ is a bijection. Consider now an arrow $f\colon X_1\to X_2$, then, for every $g\colon X_2\to G(Y)$ we have:
		\begin{align*}
			\phi_X(g)\circ F(f)&=\epsilon_Y\circ F(g)\circ F(f)\\&= \epsilon_Y\circ F(g\circ f)\\&=\phi_X(g\circ f)
		\end{align*}
		Naturality now follows at once.

		\item Let $\phi_Y\colon G(Y)\to G'(Y)$ and $\phi'_Y\colon G(Y)\to G'(Y)$ be the unique arrows making the diagram below commutative.
\[\xymatrix@C=15pt{F(G(Y)) \ar[rr]^{F(\phi_Y)} \ar[dr]_{\epsilon_Y} & &F(G'(Y)) \ar[dl]^{\epsilon'_Y} && F(G'(Y)) \ar[rr]^{F(\phi_Y)} \ar[dr]_{\epsilon'_Y} & &F(G(Y)) \ar[dl]^{\epsilon_Y}\\ & Y &&&&Y}\]
		Now, we can notice that the diagram belows commute.

	\[\xymatrix{F(G(Y)) \ar[r]^{F(\phi_Y)} \ar[dr]_{\epsilon_Y} &F(G'(Y)) \ar[r]^{F(\phi_Y)} \ar[d]_{\epsilon'_Y} & F(G(Y)) \ar[dl]^{\epsilon_Y} \\ & Y \\ F(G'(Y)) \ar[r]_{F(\phi'_Y)} \ar[ur]^{\epsilon'_Y} &F(G(Y)) \ar[r]_{F(\phi'_Y)} \ar[u]_{\epsilon_Y} & F(G'(Y)) \ar[ul]_{\epsilon'_Y}}\]	
This implies at once that $\phi_Y$ and $\phi'_Y$ are mutally inverse. \qedhere 
	\end{enumerate}

\end{proof}

\begin{lemma}\label{lem:part}
	Let $\X$ be an  an $\mathcal{M}$-adhesive category  and cosinder a pair $(m,f)$ with $m\colon M\mto X $ in $\mathcal{M}$ and $f:X\to Y$. If $(h, k)$ is a pushout complement for the pair $(m,f)$, then there exists an isomorphism $\epsilon_k:f^*(k)\to m$ in $\mathcal{M}/X$ such that $(k, \epsilon_k)$ is  a partial right adjoint at $m$ to the pullback functor $f^*\colon \mathcal{M}/Y\to \mathcal{M}/X$.
\end{lemma}
\begin{proof}
	By \Cref{prop:pbpoad}, in the diagram below the inner square is both  a pushout and a pullback, yielding the dotted isomorphism $\epsilon_k\colon f^*(Q)\to M$.
\[\xymatrix{f^*(Q) \ar@{.>}[dr]^{\epsilon_k} \ar@{>->}@/^.2cm/[drr]^{f^*(k)} \ar@/_.2cm/[ddr]_{p}\\& M \ar@{>->}[r]^{m}\ar[d]_{h}& X \ar[d]^f \\&Q\ar@{>->}[r]_{k}&Y}\]

To see that in this way we get a partial right adjoint, let $g\colon f^*(n)\to m $ be a morphism in $\mathcal{M}/X$, then we have the solid diagram below
\[\xymatrix{f^*(N) \ar[dr]^{g} \ar@{>->}@/^.2cm/[drr]^{f^*(n)} \ar[ddd]_{q}\\& M \ar@{>->}[r]^{m}\ar[d]_{h}& X \ar[d]^f \\&Q\ar@{>->}[r]_{k}&Y\\N \ar@{>->}@/_.2cm/[urr]_{n} \ar@{.>}[ur]^{t}}\]
Now, since $n$ and $f^*(n)$ are in $\mathcal{M}$we can form two other pullback squares:
\[\xymatrix{A \ar@{>->}[d]_{a}\ar@{>->}[r]^-{r}& f^*(N) \ar@{>->}[d]^{f^*(n)} & B \ar@{>->}[r]^{s} \ar@{>->}[d]_{b}& N \ar@{>->}[d]^{n}\\M \ar@{>->}[r]_{m}&X &Q \ar@{>->}[r]_k&Y\\}\]
Notice that 
\begin{align*}
	k\circ h\circ a&=f\circ m\circ a\\&=f\circ f^*(n)\circ r\\&=n\circ q \circ r
\end{align*}
Thus there exists a unique $c\colon A\to B$ as in the diagram
\[\xymatrix{A \ar@{>->}[d]_{a} \ar@{.>}[dr]^{c}\ar@{>->}[r]^{r} &f^*(N) \ar@/^.2cm/[dr]^{q} \\ M \ar@/_.2cm/[dr]_{h}&B\ar[r]^{s}  \ar@{>->}[d]_{b}& N \ar@{>->}[d]^{n}\\ &Q \ar@{>->}[r]_{k} & Y}\]
We can then consider the two diagrams below, noticing that the left one is $\mathcal{M}$-Van Kampen and in the cube on the right the front, right and left faces are pullbacks.
		\[\xymatrix@C=10pt@R=10pt{&&&&&A\ar@{>->}[dd]|\hole_(.65){a}\ar[rr]^{c} \ar@{>->}[dl]_{r} && B \ar@{>->}[dd]^{b} \ar@{>->}[dl]_{s} \\ M \ar@{>->}[dd]_{m}\ar[rr]^{h}&&Q\ar@{>->}[dd]^{k}&&f^*(N)  \ar@{>->}[dd]_{f^*(n)}\ar[rr]^(.7){q} & & N \ar@{>->}[dd]_(.3){n}\\&&&&&M\ar[rr]|\hole^(.65){h} \ar@{>->}[dl]^{m} && Q \ar@{>->}[dl]^{k} \\X\ar[rr]_{f} &&Y&&X \ar[rr]_{f} & & Y}\]
Now, we can also notice that the back square is a pullback too. To see this it is enough to apply \Cref{lem:pb1} to the diagram
\[\xymatrix{A  \ar@/^.4cm/[rr]^{a\circ r}\ar@{>->}[d]_{a} \ar[r]_{c}& \ar@{>->}[r]_{s}B \ar@{>->}[d]^{b}& N \ar[d]^{n}\\ M \ar@/_.4cm/[rr]_{f\circ m}\ar[r]^{h}& Q \ar@{>->}[r]^{k}& Y}\]

By hypothesis we have $g\colon f^*(n)\to m$ in $\mathcal{M}/X$, thus \Cref{lem:varie} yields a unique morphism $\bar{g}\colon n\to k$ in $\mathcal{M}/Y$. By the commutativity of the diagram
\[\xymatrix{f^*(N) \ar[r]^{f^*(\bar{g})}& f^*(Q) \ar[dr]^{\epsilon_k} \ar@{>->}@/^.2cm/[drr]^{f^*(k)} \ar@/_.2cm/[ddr]_{p}\\&& M \ar@{>->}[r]^{m}\ar[d]_{h}& X \ar[d]^f \\&N &Q\ar@{>->}[r]_{k}&Y}\]
we can deduce that $\epsilon_k \circ $ as wanted.
\end{proof}

We can deduce a host of properties of pushout complements from \Cref{lem:part}.

\begin{corollary}\label{lem:radj}
	Let $f\colon X\to Y$ be an arrow in an $\mathcal{M}$-adhesive category $\X$ and suppose that the left square below is a pushout while the  right one is a pullback, with $ m\colon M\to X$ and $n\colon N\to Y$ in $\mathcal{M}$.
	\[\xymatrix{M \ar@{>->}[d]_{m}\ar[r]^{h}& Q \ar@{>->}[d]^{k} & P \ar[r]^{p_1} \ar@{>->}[d]_{p_2}& N \ar@{>->}[d]^{n}\\X\ar[r]_{f}&Y &X \ar[r]_f&Y\\}\]
	Then $n\leq k$ if and only if $p_2\leq m$.
\end{corollary}
\begin{proof}  
	We start pulling back $n$ along $k$ and then pulling back the resulting arrow along $h$, getting the following two pullbacks square:
	\[\xymatrix{B \ar@{>->}[d]_{b}\ar@{>->}[r]^{s}& N \ar@{>->}[d]^{n} & A \ar[r]^{r} \ar@{>->}[d]_{a}& B \ar@{>->}[d]^{b}\\Q\ar@{>->}[r]_{k}&Y &M \ar[r]_h&Q\\}\]
	Notice that 
	\begin{align*}
		f\circ m\circ a&=k\circ h\circ a\\&=k\circ b\circ r\\&=n\circ s\circ r
	\end{align*}
	Thus there exists a unique $t\colon A\to P$ as in the diagram
	\[\xymatrix{A \ar@{>->}[d]_{a} \ar@{.>}[dr]^{t}\ar[r]^{r} &B \ar@{>->}@/^.2cm/[dr]^{s} \\ M \ar@{>->}@/_.2cm/[dr]_{m}&P\ar[r]^{p_1}  \ar@{>->}[d]_{p_2}& N \ar@{>->}[d]^{n}\\ &X \ar[r]_{f} & Y}\]
	We can then consider the cube below, in which the front, right and back faces are pullbacks.
	\[\xymatrix@C=15pt@R=15pt{&A\ar@{>->}[dd]|\hole_(.65){a}\ar[rr]^{r} \ar[dl]_{t} && B \ar@{>->}[dd]^{b} \ar[dl]_{s} \\ P  \ar@{>->}[dd]_{p_2}\ar[rr]^(.7){p_1} & & N \ar@{>->}[dd]_(.3){n}\\&M\ar[rr]|\hole^(.65){h} \ar@{>->}[dl]^{m} && Q \ar@{>->}[dl]^{k} \\X \ar[rr]_{f} & & Y}\]
	We can notice that the left square is a pullback too. To see this it is enough to apply \Cref{lem:pb1} to the diagram
	\[\xymatrix{A  \ar@/^.4cm/[rr]^{s\circ r}\ar@{>->}[d]_{a} \ar[r]_{t}& \ar[r]_{p_1} P \ar@{>->}[d]^{p_2}& N \ar[d]^{c}\\ M \ar@/_.4cm/[rr]_{k\circ h}\ar@{>->}[r]^{m}& X \ar[r]^{f}& Y}\]

@@

\begin{corollary}[Uniqueness of pushouts complements]\label{lem:pocomp} 
	contenuto...
\end{corollary}

\begin{corollary}\label{cor:pbc}
Let $\X$ be a $\mathcal{M}$-adhesive category. Given $m\colon X\mto Y$ in $\mathcal{M}$ and $f\colon Y\to Z$, let $(h_1, k_1)$ and $(h_2, k_2)$ be pushout complements of $m$  and $f$ and $W_1=\cod{h_1}$, $W_2=\cod{h_2}$. Then there exists a unique isomorphism $\phi\colon W_1\to W_2$ making the following diagram commutative.
\[\xymatrix{&X \ar@{>->}[r]^{m} \ar[d]_{h_1} \ar@/_.3cm/[ddl]_{h_2}& Y \ar[d]^{f} \\ &W_1 \ar@{.>}[dl]^{\phi} \ar@{>->}[r]_{k_1} & Z \\ W_2 \ar@{>->}@/_.3cm/[urr]_{k_2}}\]

@@

@@

\begin{lemma}\label{lem:pushandpullpb}
	Let $\X$ be a category and suppose that the diagram below is given.
	\[\xymatrix{ P \ar[r]^{p_0} \ar[d]_{p_1}& A  \ar@/^.2cm/[ddr]^{k_0}\ar[d]^{f}\\  B  \ar@/_.2cm/[drr]_{k_1} \ar[r]_g & C \ar[dr]^{h}\\&& D }\]
	Then the following hold true:
	\begin{enumerate}
		\item if the outer boundary is a pullback, then the inner square is a pullback too;
		\item if the inner square is a pullback and $h$ is mono then also the outer square is a pullback.
	\end{enumerate}
\end{lemma}
\begin{proof}\begin{enumerate}
		\item Let $e_0\colon E\to A$ and $e_1\colon E\to B$ be a pair of arrows such that 
		\[f\circ e_0=g\circ e_1\]
		Then we also have 
		\begin{align*}
		k_0\circ e_0 &= h\circ f\circ e_0\\&=h\circ g \circ e_1\\&=k_1\circ e_1
		\end{align*}
		Thus there exists a unique arrow $e\colon E\to P$ such that
		\[p_0\circ e= e_0 \qquad p_1\circ e=e_1\]
		which entails our claim.
		\item We proceed as before considering two arrows $e_0\colon E\to A$ and $e_1\colon E\to B$, but in this case we suppose that 
		\[k_0\circ e_0=k_1\circ e_1\]
		In particular, we have
		\begin{align*}
			h\circ f \circ e_0 &= k_0\circ e_0\\&=k_1\circ e_1\\&=h\circ g\circ e_1
		\end{align*}
		Thus, since $h$ is mono, we can deduce that $f\circ e_0$ and $g\circ e_1$ are equal and the thesis now follows immediately.
		\qedhere 
	\end{enumerate}
	\end{proof}

	\begin{lemma}\label{lem:cond} Let $f\colon A\to C$, $g\colon B\to C$ and $h\colon C\to D$ be arrows in a category $\X$, let also $l_0\colon A'\to A$, $l_1\colon B'\to B$ be two monomorphisms. Suppose moreover, that the following squares are pullbacks:
		\[\xymatrix@C=35pt{Q \ar[r]^{q_0} \ar[d]_{q_1} & A' \ar[d]^{h\circ f\circ l_0} & P \ar[r]^{p_0} \ar[d]_{p_1} & A \ar[d]^{f} & Q  \ar[r]^{l_0\circ q_0} \ar[d]_{l_1\circ q_1}& A \ar[d]^{h\circ f}\\ B \ar[r]_{h\circ g\circ l_1} &D& B \ar[r]_{g}& C &  B \ar[r]_{h\circ g} & D}\]
		Then there exists an arrow $\phi\colon P\to Q$ such that $(P, \{q_i\circ \phi\}_{i=0}^1)$ is a pullback for $f\circ l_0$ along $g\circ l_1$.
	\end{lemma} 
	\begin{proof}
To prove the existence of $\phi\colon P\to Q$ it's enough to consider the solid part of the diagram below.
\[\xymatrix@C=30pt{P \ar@{.>}[dr]^{\phi} \ar@/^.2cm/[drr]^{p_0} \ar@/_.2cm/[ddr]_{p_1}\\ & Q  \ar[r]^(.4){l_0\circ q_0} \ar[d]^{l_1\circ q_1}& A \ar[d]^{h\circ f} \\ &B \ar[r]_{h\circ g} & D}\]
Notice that $\phi$ is a monomorphism. If $\phi\circ a=\phi \circ b$ for some $a,b\colon T\rightrightarrows P$ then
\[\begin{split}
	p_0\circ a&=l_0\circ q_0\circ \phi \circ a\\&=l_0\circ q_0\circ \phi \circ b\\&=p_0\circ b
\end{split} \qquad \begin{split}
p_1\circ a&=l_1\circ q_1\circ \phi \circ a\\&=l_1\circ q_1\circ \phi \circ b\\&=p_1\circ b
\end{split} \]
